|

### 2020-07-14 paper.Rmd
Start rewriting implementation.

We therefore allow trimming of the projected observations, using $R$ as a free parameter of the display function. When selecting a value $R$ that is smaller than the maximum distance from the center, we need to ensure that the projected radius of points is always smaller than $R$ before rescaling according to Eq. \ref{eq:resc}. We do this by replacing $r_i$ as
\begin{equation}
\tilde{r}_i = \min(r_i, R)
\end{equation}
for each observation $i = 1, ..., n$.

@@

# Implementation {#sec:implementation}

XXX this should explain how the display integrates with the tour, in terms of a data pipeline: p-D data -> projection matrix from the tour -> d-D projected data -> radial transformation -> final display

XXX include short tour intro here with relevant references

@@

The sage display has been implemented in R [@rref] in the `display_sage` method in the `tourr` package [@tourr]. This display has two important new input parameters, the scaling `s` for computing $p_{eff}$, and the overall scale `R`. Both these parameters are described in Section \ref{sec:params} above.

>>>>

We therefore allow trimming of the projected observations, using $R$ as a free parameter of the display function. When selecting a value $R$ that is smaller than the maximum distance from the center, we need to ensure that the projected radius of points is always smaller than $R$ before rescaling according to Eq. \ref{eq:resc}. We do this by replacing $r_i$ as
\begin{equation}
\tilde{r}_i = \min(r_i, R)
\label{eq:cutR}
\end{equation}
for each observation $i = 1, ..., n$.

@@

## Implementation as a dynamic display {#sec:implementation}

While the radial transformation can in general be used with any low-dimensional display that suffers from data piling, it is most useful when combined with a dynamic display showing a sequence of interpolated low-dimensional projections obtained when running a tour. We have implemented it as a new display method in the `display_sage` method in the `tourr` package [@tourr] in R [@rref].

We can think of the display functions as part of a data pipline obtained when running a tour. The initial step is pre-processing the data ($\mathbf{X}$, an $n \times p$ matrix), typically this includes centering and scaling using either the overall range or the variance. The tour is then looping over the following steps:

1. Obtain projection matrix $\mathbf{A}$. For $d$-dimensional projections this is an orthonormal $p \times d$ matrix. To ensure the smooth rotation of projections, each new $\mathbf{A}$ is obtained as an interpolated step in the sequence, as explained in @BCAH05.
2. Project the data by computing $\mathbf{Y} = \mathbf{X}\cdot\mathbf{A}$.
3. Map $\mathbf{Y}$ to the display to re-draw the projected data. We can think of the mapping as a function $f(\mathbf{Y})$, and for $d=2$ this typically maps the projected points onto a scatter plot display. Note that in general the mapping may also depend on additional information, for example @sliceTour is using the distance from the projection plane to show a slice display.

These three steps are available in the `tourr` package, which includes different options for basis selection, and different display functions. Since the implementation is modular, we can add the sage display as a new function on the projected data, $f(\mathbf{Y})$. This mapping is done in the following steps:

1. Center the 2-dimensional matrix $\mathbf{Y}$ and compute its polar coordinate representation $(r, \theta)$.
2. For each observation, first use Eq. \ref{eq:cutR} to ensure the radius $r$ is within the specified range, and then apply the radial transformation defined in Eq. \ref{eq:resc} to obtain $r'$.
3. Use the transformed radial coordinate $r'$ to re-compute the mapping onto Euclidean coordinates $(x, y)$.
4. To fit the final projection onto the plotting canvas ranging between $[-1,1]$, we rescale each observation by $0.9 \cdot \mathrm{half\_range}$.

@@

 This display has two important new input parameters, the scaling `s` for computing $p_{eff}$, and the overall scale `R`. Both these parameters are described in Section \ref{sec:params} above.
